---
author: "Han Zhang"
output: html_document
---

```{r setup, include=FALSE}
library(quanteda)
library(Rtsne)

library(data.table) # this package help you to read large files quicker
reviews <- fread("../lec4-text-basics-codes/train.csv") # fread() is a function in data.table
reviews$label <- reviews$label # coerce 0/1 to category
# text to corpus
twcorpus <- corpus(reviews$text)
# the below commented lien won't work; read the help file
# doc.term <- dfm(twcorpus,  ngrams=1:3, tolower=TRUE, stem=TRUE, remove_punct = TRUE, remove_url=TRUE,  verbose=TRUE, remove=c(stopwords("english"), "the", "t.co", "https", "rt", "amp", "http", "t.c", "can", "u", "<", ">", "br"))

# transform corpus to tokens
tokens <-  tokens(twcorpus, remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE, remove_url = TRUE)
tokens <- tokens_remove(tokens, c(stopwords("english"), "the", "t.co", "https", "rt", "amp", "http", "t.c", "can", "u", "<", ">", "br", " ")) 
tokens <- tokens_wordstem(tokens)
tokens <- tokens_ngrams(tokens, n = 1, concatenator = ".")

# then generate document-
doc.term <- dfm(tokens,  verbose=TRUE)
print (dim(doc.term))

# and discard too infrequent words; 
doc.term <- dfm_trim(doc.term, min_docfreq=500, verbose=TRUE)
print (dim(doc.term))
```

```{r}
X <- as.matrix(doc.term)
```


## pca

The native implementation is really slow..

You have to search for quicker alternatives. This is something you should try if your algorithm is too slow: just google XXX algorithm + fast implementation.
Sometimes there will be faster versions; sometimes not.


```{r}
# r native function prcomp() is very slow
# pc <- prcomp(X, scale = FALSE)
```


## select first 1500 documents
- because we want to get vectors, but now each column is a word vector, we transpose it
- now each row is a word vector
```{r}
library(gmodels)

X_subset = t(X[1:1500, ])

pc <- fast.prcomp(X_subset, retx = TRUE, center = TRUE, scale. = FALSE, tol = NULL)

```

```{r}
# you can take first several principal components in this way
pca_vector <- pc$x[,1:2]
plot(pca_vector)
pca_vector[pca_vector[,1]>30,]
```

tsne


```{r}
tsne <- Rtsne(X_subset, dims = 2, partial_pca = T, perplexity=30, verbose=TRUE, max_iter = 50)

embedding <- tsne$Y
plot(embedding)

mx1 <-which.max(embedding[,1])
embedding[mx1,]
colnames(X)[mx1]

mx2 <-which.min(embedding[,2])
embedding[mx2,]
colnames(X)[mx2]
```
